---
title: "knitr_presentations"
author: "Robert Olendorf"
date: "March 1, 2016"
output: html_document
---

```{r, setup, echo=FALSE, results='hide', warning=FALSE, message=FALSE, error=TRUE}
  if(!require("pander"))
  {
    install.packages("pander", repos="http://cran.rstudio.com/", dependencies = TRUE)
  }
  
  library("pander", lib.loc="~/Library/R/3.2/library")

  framework.data <- read.csv('frameworks.csv')

```

## Knitr and R presentations

Knitr is also a great way to make slide presentations. Knitr uses Javascript libraries like *Reveal.js* or *Deck.js* to power the slides, so if you have used those this may be familiar to you. It's not always better than PowerPoint or similar options. I like Knitr for technical talks, especially with lots of charts and data. Also they are great if you want to publish your slides on the web. Powerpoint is great when you want slides fast. Also, its easier to style and format slides in Powerpoint. Knitr takes a fair amount of HTML and CSS to create highly formatted slides.

Knitr also gives us a number of methods for creating slides. The first is to choose the **_Presentation_** option when you choose to create a new Markdown file. This gives you three options. **_HTML (ioslides)_**, **_HTML (Slidy)_** and **_PDF (Beamer)_**. The first two use different Javascript frameworks to create HTML slides. The last option will convert the Markdown into LaTeX then PDF (I think). 

The second method is to choose an **_R Presentation_**. This creates a *.Rpres* file thats essentially Markdown again. 

The final option and the one we'll use, is to use the **_Slidify_** package. It's a little more complicated, but it also gives us more options for layouts and theming. It essentially provides an interface for several well known Javascript slide frameworks. 

## Slidify

### Setup

First we need to load the packages. **_Slidify_**. **_Slidify_** has pretty good documentation at [slidify.org](http://slidify.org/), but its a little out of date so use the [Slidify Github](https://github.com/ramnathv/slidify) pages instead. To install the packages, first make sure the **_devtools_** package is installed. Go to the **Packages** tab on the lower right and see if **_devtools_** is listed. If it is, make sure its checked. If its not listed, click the **Install** button and enter **_devtools_** in the search box. Make sure *install dependencies* is checked and click *install*.

To install **_Slidify__** we need to download two packages from Github using the following lines on the console.

```r
   install_github("ramnathv/slidify")
   install_github("ramnathv/slidifyLibraries")
   library('slidify)
   library('slidifyLibraries')
```

Next we create a stub slide deck by running the following command on the console.

```r
  author('mtcars_presentation')
```

The result will be a markdown document that looks somethign like this.

```markdown
  ---
  title       : 
  subtitle    : 
  author      : 
  job         : 
  framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
  highlighter : highlight.js  # {highlight.js, prettify, highlight}
  hitheme     : tomorrow      # 
  widgets     : []            # {mathjax, quiz, bootstrap}
  mode        : selfcontained # {standalone, draft}
  knit        : slidify::knit2slides
  ---
  
  ## Read-And-Delete
  
  1. Edit YAML front matter
  2. Write using R Markdown
  3. Use an empty line followed by three dashes to separate slides!
  
  --- .class #id 
  
  ## Slide 2
  
  


```
   
The first four lines are pretty self explanatory and are used by some frameworks to create a title slide for you. Go ahead and fill it in with your info. Next we can choose the framework. There are several to choose from listed below, some of them also have multiple themes.

```{r, framework-table, echo=FALSE}
  pander(framework.data)
```

Try some of the frameworks out. And try setting the theme as well. To set a theme, adjust your YAML front matter to look like this.

```r
  ---
  title       : Motor Trend Cars Analysis
  subtitle    : Lets Talk About Cars
  author      : Robert Olendorf
  job         : Gadabout
  framework   : deckjs        # {io2012, html5slides, shower, dzslides, ...}
  deckjs:
    theme: neon
  highlighter : highlight.js  # {highlight.js, prettify, highlight}
  hitheme     : tomorrow      # 
  widgets     : []            # {mathjax, quiz, bootstrap}
  mode        : selfcontained # {standalone, draft}
  knit        : slidify::knit2slides
  ---
```

You will notice different frameworks have different ideas about how to display a slide show. Some have title slides, some not. Some show all the slides at once. Some frameworks have nice transitions, and clever navigation too. We'll spend our time with **_Deck.js_** only because I know it best. You can also try different code highlighter and themes. The widgets we'll get to later. They add functionality and interactivity to your slides.

The last two options you shouldn't have to change. The mode determines if the slide application is selfcontained (packages all the dependences and ships them together) or allows it to make calls to external resources. Usually selfcontained is the best option. The Knit option just tells Knitr what to do.





Then make your second slide look like this. Note. You'll need an image of a car. You can grab mine from Github or get your own. Make sure its a PNG or alter your code accordingly.

### Adding Slides

Our first slide can just be bullet points.

```r
  ---
  
  ## Some Fun Facts About cars
  
 
  ### Facts
  
  1. Cars have four wheels
  2. Some cars are really fast
  3. The Gemlin is very underrated
  
  ### Fictions
  
  1. Cars can drive them selves
  2. We'll have flying cars in 2010
  
  ---

```

Knitr and Slidify do pretty well at organizing the material. Depending on the framework, the first heading is treated as a Deck Title, the second heading is a slide title, and other content is dealt with accordingly.

We can add some actual slides now. With **_Deck.js_**, we need to add our own title slide (I think). Lets do that. EAch slides ends with three dashes `---`.
 

Add the following first. It can be on this slide, or in the first. Needs to be before we use **_png_**.

We can add slide with an image. A car seems appropriate. We are also making use of some R packages (**_png_** and **_grid_**) to make image processing and resizing possible. We need to add the conditional install of **_png_** since it doesn't come in the default packages. We also need to require both.

```r

  if(!require("png"))
  {
    install.packages("png", repos="http://cran.rstudio.com/", dependencies = TRUE)
  }
  library(png) 
  
```


```r

## Cars Look Really nice

  `` `` ``{r, a cool ratrod, fig.width=8.0, echo=FALSE, fig.align='center'}

   
  library(grid)
  img <- readPNG("assets/img/ratrod.png")
   grid.raster(img, just="centre")

  `` `` ``

---
```

We can add a table too. This took some research as the original **_Pander_** table didn't work in **_Slidify_**. After some searching on Stack Overflow, I found this answer http://stackoverflow.com/questions/18810749/how-to-print-tables-in-slidify).

```r

  ## Another Peak At Our Data
  
  `` `` ``{r, mtcars-data, echo=FALSE}
     car.data <- head(mtcars[, 1:6], 5)
     set.alignment(row.names = 'left')
     pander(car.data, style = 'rmarkdown', split.cells = Inf, caption = "The first nine columns of mtcars")
  `` `` ``
  
  ---
```

I've also truncated the rows and columns even more to make it fit on the slide better. If you are using **_Deck.js_**, it is pretty bare. **_Io2012_** looks nicer. This can all be changed with some CSS. We can dress up the table a little now. 

Adding an analysis created some problems. I finally got a nice table using **_xtable_**. 

```r
  ---

  ## Analyzing The Data

  `` `` ``{r, mtcars-regression, results='asis'}
     
    library(xtable)
    print(xtable(summary(lm(mtcars$mpg ~ mtcars$wt))),type='html')
  `` `` ``

  ---
```

### Slide Layouts

Let's make a two column slide, the regression results on one line and a plot on the other. This requires changing our slide layout. Some frameworks have that layout built-in. If your using Deckjs you have to make your own layout. You can see what layouts come with a particular frame work by looking in ` libraries\frameworks\<framework_name>\layouts `. You can specify an existing layout by putting ` &layout_file_name ` after the three dashes that end and start a slide. Knitr will look for that file in ` libraries\frameworks\<framework_name>\layouts ` or in ` assets\layouts `. The latter being where you put your custom layouts. If you have a layout in your local assets with the same name as a layout in the framework's layouts, your version gets used.

Lets make a custom layout for **_deckjs_**. I pretty much just "borrowed" the code from *io2012* layouts for two column and came up with this. Save it in ` assets\layouts ` as `twocol.html`.

```html
  ---
  layout: slide
  ---
  <p> foo </p>
  {{{ slide.content }}}
  <hr>
  <div style='float:left;width:48%;' class='centered vert-align two-col'>
    {{{ slide.left.html }}}
  </div>
  <div style='float:right;width:48%;' class='centered vert-align two-col'>
    {{{ slide.right.html }}}
  </div>
```

I added some classes to allow us to fix some styling with CSS later. Then in your slides make the following slide.

```r
  ---&twocol
  
  ## Analyzing The Data
  
  ### We can display two columns with a custom layout.
  
  *** {name: left}
  
  
  `` `` ``{r, mtcars-regression, results='asis', echo=FALSE}
     
    library(xtable)
    print(xtable(summary(lm(mtcars$mpg ~ mtcars$wt))),type='html')
  `` `` ``
  
  *** {name: right}
  
  
  `` `` ``{r, mtcars-plot, echo=FALSE}
    plot(mtcars$wt, mtcars$mpg)
  `` `` ``
---
```

The result works but its ugly. We need to style the table, and see if we can vertically align. We can add custom *.css* in our ` assets\css ` directory. Slidify automatically reads and uses any *.css* file in there. Create a file and save it as `custom.css` in your `assets\css` directory. With some trial and error, I improved the display a bit with this.

```html
  .deck-container table {
  	vertical-align: middle;
  	height: 100%;
    margin-right: auto;
    margin-left: auto;
  }
  
  .deck-container table, .deck-container th, .deck-container td {
    border: 1px solid black;
    padding: 10px;
  }

```

You can also add classes and IDs to your slides. Its very easy. Just do somethign like this.

```r
  --- .fun-facts #about-cars
  
  ## Some Fun Facts About cars
  
  ### Facts
  
  1. Cars have four wheels
  2. Some cars are really fast
  3. The Gemlin is very underrated
  
  ### Fictions
  
  1. Cars can drive them selves
  2. We'll have flying cars in 2010
  
  ---
  
```

The html will look somethign like this. Each slide is a `<section>` and the class and ID get inserted there.

```html
  <section class="slide fun-facts" id="about-cars">
    <h2>Some Fun Facts About cars</h2>
    <h3>Facts</h3>
  
    <ol>
      <li>Cars have four wheels</li>
      <li>Some cars are really fast</li>
      <li>The Gemlin is very underrated</li>
    </ol>
    
    <h3>Fictions</h3>
    
    <ol>
      <li>Cars can drive them selves</li>
      <li>We&#39;ll have flying cars in 2010</li>
    </ol>
  
  </section>
```

We'll look a bit more of slides in the when we explore Shiny. We can make slides more interactive there.



